---
title: "Engagement and Attendance Survey Report"
author: "Chris Martin"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    toc: true
    theme: readable
    code_folding: hide
    css: mystyle.css
---

Department of Psychology, University of Sheffield, July 2023

In June and July 2023 we carried out a survey of students enrolled on our undergraduate BSc Psychology programme to try to better understand student perceptions of course engagement and attendance. Approval for the research was provided by the Ethics Committee of the University of Sheffield.

This is a report on the results of that survey, produced using R Markdown.

This report contains preliminary analysis only. A full write up is in progress and we plan to submit the work for publication.

## Prepare environment and load data

Make the the qualtRics, likert and tidyverse libraries available and then load the data.
Data is combined across the Level 1 and Level 2 surveys into a single data frame.
```{r, message=FALSE}
library('qualtRics')
library('likert')
library('tidyverse')

L1data <- data.frame(read_survey("Psychology Student Engagement Survey 2023_20 July 2023_10.51.csv")) #load Level 1 data
L2data <- data.frame(read_survey("Level 2 Psychology Student Engagement Survey 2023_20 July 2023_15.30.csv")) #load Level 1 data
EngQs <- read.csv2("EngQs.csv", sep = ",") #load column names
Data <- rbind(L1data, L2data) #combine data
```

## Screen and clean data

Order data by Qualtrics 'progress' measure and inspect rows (rows 20-26 reveal relevant cut-off point, as below).

```{r}
Data <- arrange(Data, Progress)
y<-data.frame(lapply(Data[c(20:26),c(70:80)], substr, 1, 3))
y
```

Up to and including row 22, there are missing values. Therefore, remove first 22 rows. Then apply longer column names to facilitate Likert data visualisation later.

```{r}
DataComplete <- Data[c(23:162),]
colnames(DataComplete)[c(5, 12:92)]<-EngQs[c(5, 12:92),3]
```

Next, show duration data (how long it took respondents to complete the survey).

```{r}
ggplot()+
  geom_boxplot(aes(y=DataComplete$Duration))+
  geom_jitter()+
  scale_x_discrete()+
  geom_boxplot(outlier.shape = NA)+
  ylab("Time to complete survey (hrs)")+xlab("All Respondents")
```

Two entries with very long duration: inspection of responses indicates they are complete and fine to include.

## Demographics and other features

> Basic piechart for level of study

```{r}
ggplot(data.frame(table(DataComplete$Level)), aes(x="", y=Freq, fill=Var1))+
  geom_bar(stat="identity", width=1) +
  geom_text(aes(label = round(100*Freq/sum(Freq), digits=1)), position = position_stack(vjust = 0.5)) +
  coord_polar("y", start=0) +
  labs(title = "Level of Study", fill = "Level")+
  theme_void()+
  theme(plot.title = element_text(hjust = 0.5))
```

> Basic piechart for language

```{r}
ggplot(data.frame(table(DataComplete$Language)), aes(x="", y=Freq, fill=Var1))+
  geom_bar(stat="identity", width=1) +
  geom_text(aes(label = round(100*Freq/sum(Freq), digits=1)), position = position_stack(vjust = 0.5)) +
  coord_polar("y", start=0) +
  labs(title = "Is English your first language?", fill = "Response")+
  theme_void()+
  theme(plot.title = element_text(hjust = 0.5))
```

> Basic piechart for ethnicity

``` {r}
ggplot(data.frame(table(DataComplete$Ethnicity)), aes(x="", y=Freq, fill=Var1))+
  geom_bar(stat="identity", width=1) +
  geom_text(aes(label = round(100*Freq/sum(Freq), digits=1)), position = position_stack(vjust = 0.5)) +
  coord_polar("y", start=0) +
  labs(title = "Ethnicity", fill = "Response")+
  theme_void()+
  theme(plot.title = element_text(hjust = 0.5))
```

> Basic piechart for gender identity

```{r}
ggplot(data.frame(table(DataComplete$Gender)), aes(x="", y=Freq, fill=Var1))+
  geom_bar(stat="identity", width=1) +
  geom_text(aes(label = round(100*Freq/sum(Freq), digits=1)), position = position_stack(vjust = 0.5)) +
  coord_polar("y", start=0) +
  labs(title = "Gender Identity", fill = "Response")+
  theme_void()+
  theme(plot.title = element_text(hjust = 0.5))
```

> Basic piechart for traveltime

```{r}
ggplot(data.frame(table(DataComplete$Travel_Time)), aes(x="", y=Freq, fill=Var1))+
  geom_bar(stat="identity", width=1) +
  geom_text(aes(label = round(100*Freq/sum(Freq), digits=1)), position = position_stack(vjust = 0.5)) +
  coord_polar("y", start=0) +
  labs(title = "Travel Time to University", fill = "Response")+
  theme_void()+
  theme(plot.title = element_text(hjust = 0.5))
```

> Basic piechart for caring responsibilities

```{r}
ggplot(data.frame(table(DataComplete$Caring)), aes(x="", y=Freq, fill=Var1))+
  geom_bar(stat="identity", width=1) +
  geom_text(aes(label = round(100*Freq/sum(Freq), digits=1)), position = position_stack(vjust = 0.5)) +
  coord_polar("y", start=0) +
  labs(title = "Caring Responsibilities", fill = "Response")+
  theme_void()+
  theme(plot.title = element_text(hjust = 0.5))
```

## Plot Likert data for engagement perceptions

```{r, out.width="150%"}

#define factor levels for likert
liklev <- c("Strongly disagree", "Disagree", "Neither agree nor disagree", "Agree", "Strongly Agree")

#set factor and levels for engagement perceptions questions
for (x in 22:46) {
  DataComplete[,c(x)] <- factor(DataComplete[,c(x)], levels = liklev)
}

# plot(likert) perceptions of engagement
plot(likert(DataComplete[c(22:46)]), legend.position="right", text.size=2) +
  theme(axis.text=element_text(size=5)) +
  theme(panel.grid = element_line(color = "gray90"))

```

## Plot Likert data for enablers/barriers

```{r, out.width="150%"}

#set factor and levels for enablers/barriers questions
for (x in 60:85) {
  DataComplete[,c(x)] <- factor(DataComplete[,c(x)], levels = liklev)
}
  
# plot(likert) enablers/barriers to attendance
plot(likert(DataComplete[c(60:85)]), legend.position="right", text.size=2) +
  theme(axis.text=element_text(size=5)) +
  theme(panel.grid = element_line(color = "gray90"))
```






